---
title: "STATS506-HW3-JunjieZeng"
author: "Junjie Zeng"
format: html
editor: visual
---

# STATS506 Problem Set #2

## Repository

https://github.com/JunjieZeng1/STATS-506.git

## Problem 1 Dice Game

### a

```{r, warning = False}
# Load packages
library(haven)
library(dplyr)
library(knitr)
library(tidyr)
library(broom)
library(pscl)
library(DescTools)
```

```{r}
# Load data
vix_d <- read_xpt("/Users/junjiezeng/Desktop/git/STATS506/HW3/VIX_D.XPT")
demo_d <- read_xpt("/Users/junjiezeng/Desktop/git/STATS506/HW3/DEMO_D.XPT")
```

```{r}
# Merging data
df <- data.frame(inner_join(vix_d, demo_d, by = "SEQN"))
head(df)
```

```{r}
# Check sample size is 6980
nrow(df)
```

### b

```{r}
# Create age categories
df <- df |> 
  mutate(age_levels = cut(RIDAGEYR, breaks = seq(0, 90, by = 10), right = FALSE, 
                          labels = c("0-9", "10-19", "20-29", "30-39", "40-49", 
                                      "50-59", "60-69", "70-79", "80+")))
```

```{r}
glass_prop <- df |> 
  select(c(VIQ220, age_levels)) |> 
  filter(VIQ220 %in% c(1,2)) |> 
  group_by(age_levels) |> 
  summarize(glass = sum(VIQ220 == 1), 
            total = n(), 
            proportion = sum(VIQ220 == 1) / total)
```

```{r}
# Create a table using kable
kable(glass_prop, 
      format = "html", 
      caption = "Proportion of Wearing Glasses/Contact Lenses", 
      col.names = c("Age", "Wearing Glasses/Contacts", "Total", "Proportion"),
      digits = 2)
```

### c

```{r}
# Clean the data
df_cleaned_for_pc <- df |> 
  select(c(VIQ220, INDFMPIR, RIDAGEYR, RIAGENDR, RIDRETH1)) |> 
  drop_na(VIQ220) |> 
  mutate(is_glass = ifelse(VIQ220 == 1, 1, 0), 
         gender = as.factor(RIAGENDR), 
         race = as.factor(RIDRETH1), 
         age = RIDAGEYR, 
         PIR = INDFMPIR)
```

```{r}
# Model 1
model1 <- glm(is_glass ~ age, data = df_cleaned_for_pc, family = "binomial")

# Model 2
model2 <- glm(is_glass ~ age + race + gender, data = df_cleaned_for_pc, family = "binomial")

# Model 3
model3 <- glm(is_glass ~ age + race + gender + PIR, data = df_cleaned_for_pc, family = "binomial")
```

```{r}

model_summary <- function(lg) {
  odds_ratio <- exp(coef(lg))
  pseudo_r2 <- PseudoR2(lg)["McFadden"]
  aic <- lg$aic
  sample_size <- nrow(lg$model)
  
  return(list(odds_ratio = odds_ratio, 
              pseudo_r2 = pseudo_r2, 
              AIC = aic, 
              sample_size = sample_size))
}


model1_summary <- model_summary(model1)
model2_summary <- model_summary(model2)
model3_summary <- model_summary(model3)
```

```{r}
table_models <- data.frame(
  Model = c("Age", 
            "Age, Race, Gender", 
            "Age, Race, Gender, PIR"),
  
  Odds_Ratio_Intercept = c(round(model1_summary$odds_ratio[1], 2),
                           round(model2_summary$odds_ratio[1], 2),
                           round(model3_summary$odds_ratio[1], 2)),
  
  Odds_Ratio_Age = c(round(model1_summary$odds_ratio[2], 2),
                     round(model2_summary$odds_ratio[2], 2),
                     round(model3_summary$odds_ratio[2], 2)),
  
  Odds_Ratio_Female = c(round(model1_summary$odds_ratio[7], 2),
                        round(model2_summary$odds_ratio[7], 2),
                        round(model3_summary$odds_ratio[7], 2)),
  
  Odds_Ratio_Other_Hispanic = c(round(model1_summary$odds_ratio[3], 2),
                                round(model2_summary$odds_ratio[3], 2),
                                round(model3_summary$odds_ratio[3], 2)),
    
  Odds_Ratio_Non_Hispanic_White = c(round(model1_summary$odds_ratio[4], 2),
                                    round(model2_summary$odds_ratio[4], 2),
                                    round(model3_summary$odds_ratio[4], 2)),
  
  Odds_Ratio_Non_Hispanic_Black = c(round(model1_summary$odds_ratio[5], 2),
                                    round(model2_summary$odds_ratio[5], 2),
                                    round(model3_summary$odds_ratio[5], 2)),
  
  Odds_Ratio_Other_Race = c(round(model1_summary$odds_ratio[6], 2),
                                  round(model2_summary$odds_ratio[6], 2),
                                  round(model3_summary$odds_ratio[6], 2)),
  
  Odds_Ratio_PIR = c(round(model1_summary$odds_ratio[8], 2),
                     round(model2_summary$odds_ratio[8], 2),
                     round(model3_summary$odds_ratio[8], 2)),
  
  Sample_Size = c(model1_summary$sample_size, 
                  model2_summary$sample_size, 
                  model3_summary$sample_size),
  
  Pseudo_R2 = c(round(model1_summary$pseudo_r2, 4),
                round(model2_summary$pseudo_r2, 4),
                round(model3_summary$pseudo_r2, 4)),
  
  AIC = c(model1_summary$AIC, 
          model2_summary$AIC, 
          model3_summary$AIC)
)
```

```{r}
kable(table_models, 
      format = "html", 
      col.names = c("Model", 
                    "Odds Ratio Intercept", 
                    "Odds Ratio Age",
                    "Odds Ratio Female",
                    "Odds Ratio Other Hispanic",
                    "Odds Ratio Non-Hispanic White", 
                    "Odds Ratio Non-Hispanic Black",
                    "Odds Ratio Other Race",
                    "Odds Ratio PIR",
                    "Sample Size", 
                    "Pseudo R-squared", 
                    "AIC"))
```

### d

```{r}
# Summary model 3
summary(model3)
```

To test whether the odds of men and women being wears of glasess/contact lenses for distance vision differs, we only need to look at the summary table of the model 3. On the gender2 (female) row, a hypothesis test was automatically done by R telling us whether there is a difference between the odds. As the table shown, the z-value is 9.487, and the p-value is \< 2e-16, which is significantly less than the significance level 0.05, rejecting the null hypothesis that the odds of men and women being wears of glasess/contact lenses for distance vision do not differ. We have enough evidence to conclude that they differ.

To test whether the proportion of wearers of glasses/contact lenses for distance vision differs between men and women, we shall use a chi-squared test.

```{r}
# Perform chi-squared test
chisq_test <- chisq.test(table(df_cleaned_for_pc$gender, df_cleaned_for_pc$is_glass))
chisq_test
```

From the summary above, the test statistic is 70.593 and the degree of freedom is 1. The p-value for the Chi-squared test is \< 2.2e-16. So we reject the null hypothesis that the proportions of men and women being wears of glasess/contact lenses for distance vision do not differ. We have enough evidence to conclude that the proportions differ.

### Reference for Problem 1

#### Calculate odds ratio

https://stackoverflow.com/questions/41384075/r-calculate-and-interpret-odds-ratio-in-logistic-regression \#### Calculate pseudo-R\^2 https://search.r-project.org/CRAN/refmans/DescTools/html/PseudoR2.html


## Problem 2

### a
```{r}
# Packages
library(DBI)
library(RSQLite)
```

```{r}
# Import the SQLite database of the sakila data
sakila <- dbConnect(RSQLite::SQLite(), "/Users/junjiezeng/Desktop/git/STATS506/HW3/sakila_master.db")
sakila
```
```{r}
dbListTables(sakila)
dbListFields(sakila, 'film')
```

```{r}
dbGetQuery(sakila, "
SELECT release_year, COUNT(*) AS Num_films
  FROM film
  WHERE release_year = (SELECT MIN(release_year) FROM film)
")
```

### b
```{r}
library(tidyverse)
```

```{r}
# Method 1
df <- dbGetQuery(sakila, '
SELECT f.film_id, fcc.category_id, fcc.name
  FROM film AS f
      INNER JOIN
      (SELECT c.category_id, c.name, fc.film_id
         FROM category as c
              INNER JOIN film_category AS fc ON c.category_id = fc.category_id) AS fcc ON f.film_id = fcc.film_id
')

df |> 
  group_by(name) |> 
  summarize(total = n()) |> 
  arrange(total)
```
Music movie is the least common in the data, there are 51 movies of this genre. 


```{r}
# Method 2
dbGetQuery(sakila, '
SELECT f.film_id, fcc.category_id, fcc.name, COUNT(*) AS count
  FROM film AS f
      INNER JOIN
      (SELECT c.category_id, c.name, fc.film_id
         FROM category as c
              INNER JOIN film_category AS fc ON c.category_id = fc.category_id) AS fcc ON f.film_id = fcc.film_id
  GROUP BY fcc.name
  ORDER BY count
  LIMIT 1
')
```

Same result as the method 1. Music movie is the least common in the data, there are 51 movies of this genre.    


### c

```{r}
# Method 1
df2 <- dbGetQuery(sakila, '
SELECT cocia.country
  FROM customer AS c
    INNER JOIN 
    (SELECT a.address_id, co.country
       FROM address AS a
         INNER JOIN city AS ci ON a.city_id = ci.city_id
         INNER JOIN country AS co ON ci.country_id = co.country_id) AS cocia
     ON cocia.address_id = c.address_id
')

df2 |> 
  group_by(country) |> 
  summarize(total = n()) |> 
  filter(total == 13)
```
Argentina and Nigeria are countries have exactly 13 customers. 

```{r}
# Method 2
dbGetQuery(sakila, '
SELECT cocia.country, COUNT(*) AS customer_count
  FROM customer AS c
    INNER JOIN 
    (SELECT a.address_id, co.country
       FROM address AS a
         INNER JOIN city AS ci ON a.city_id = ci.city_id
         INNER JOIN country AS co ON ci.country_id = co.country_id) AS cocia
     ON cocia.address_id = c.address_id
  GROUP BY cocia.country
  HAVING COUNT(*) = 13
')
```
Same as the method 1, Argentina and Nigeria are countries have exactly 13 customers.